### 2015-08-13 reviews/response-rev1.txt
added reviews and response

Dear Dina, dear reviewers:

thank you very much for your time and good feedback!

We revised our paper according to your suggestions.

In the following, we discuss the most important changes.

We hope that the new version meets your expectations, and are happy to take any further feedback.

With kind regards,

Liron, Petr, Stefan