### 2015-08-25 reviews/response-rev1.txt
rev plan

Dear Dina, dear reviewers:

thank you very much for your time and good feedback!

We revised our paper according to your suggestions.

In the following, we discuss the most important changes.

We hope that the new version meets your expectations, and are happy to take any further feedback.

With kind regards,

Liron, Petr, Stefan

Reviewer 1
==========

> Overall, the work seems useful and correct.
> My main concern is scope.
> The paper is essentially a note on OpenFlow hacking, and thus might be
> of limited interest to many CCR readers. However, OpenFlow and similar
> SDN approaches have seen sufficient adoption and research interest in
> recent years that there is utility in this paper appearing in a venue with
> broad reach.

FIXME: discuss more broadly the interface and meta data approach and examples?

> A lesser concern is novelty. With the addition of atomic bundles to
> OpenFlow, it's not completely surprising that the atomic CAS operation
> can be implemented, but there is definite value in having it spelled out
> cleanly.

FIXME: it is non trivial and inband perspective novel

> I found the brief argument for inband
> control vs outofband control unconvincing. There are many highperformance
> systems that use consensus, and a full architectural
> comparison of the two approaches to
> SDN is beyond the scope of this paper. I certainly believe that inband
> control is a useful and interesting design choice, but not just because of
> FLP. It also seems almost certain that a full system would require outofband
> coordination between controllers: a full transactional system needs
> application specific ways of resolving conflicts, which will require
> coordination.

FIXME: right, that is what we are currently exploring also in related work

> I also found 2nd example of the limitations of bundles confusing. A
> simple concrete example would be greatly helpful, and strength the
> argument.

FIXME

> Typos/corrections:
> p.1 col. 2 para. (Contributions): 2nd "conflicts" => "a conflict"

FIXME

> "does not affect" => "affects neither"

FIXME

> " " para. 3: "promote" => "propose" (?)

FIXME

> p. 3 col. 2 para (Configuration space): "provided a total addressing scheme" sentence is confusing.

FIXME

Reviewer 2
==========

> The idea of the paper is really nice and I believe it can be used in
> practice. However, the presentation of the paper needs major work it
> feels to be switching between highlevel
> overview versus lowlevel details is
> background section necessary? Cannot you introduce details like
> bundles, overlap_checking and actual OpenFlow flags progressively and
> on examples?
> maybe start with another example with controllers updating
> overlapping flows. Then explain ability to do overlap_check. Then
> introduce loadbalancing example and the need for external
> synchronization. Transactions/bundles are kindof
> orthogonal and could
> be explained after you state that you want executeatomic{ops} and
> that there is a native support for it in OpenFlow if ops are OF commands.
> Figure 1 time axis is too inconspicous
> when you described execute{op1,...} and executeatomic{
> op1, ...} I was totally confused what is the definition good for. Only later on I
> realized that operations there are actually not OpenFlow operations but
> your custom synchronization primitives. Please, clarify that in the text.
> what does "mask=1...1" mean? Is it a) exact match or b) everything
> wildcarded? From analysing algorithm 5 I believe it is a)

> Do the algorithms 24 really need to use value=x.x?
> Cannot it be "0..0 . x" ? Or "1...1 . x"? Or for an arbitrary fixed
> constant "c.x" ? Maybe elaborate in the text on that point because it
> looks like magic to me.
> In any case, you should explain more why "x masked by self1"
> overlaps with "x masked by self2" exactly iff self1 != self2
> On my first reading I missed that claim() operation can actually allow
> more controllers to claim the same value. Please clarify in the text.
> What is the value of such operation? Actually, it looks like a kindof
> "read lock". following the thought, additionally to claim() and unclaim(), could you
> have algorithms for exclusive lock() and unlock()? Maybe even explain it
> sooner than claim() so that people can think about wellknown
> semantic first.
> There looks to be an inconsistency of what happens with check_overlap
> flag for the exactly same rule:
> You say "a FlowMod command can be equipped with a the check
> overlap flag: if the switch maintains a flow entry with an overlapping but
> not identical match part with the same priority but a different action,
> and if the two rules have the same priority, then FlowMod will fail."
> OF 1.4 spec says "For add requests (OFPFC_ADD) with the
> OFPFF_CHECK_OVERLAP flag set, the switch must first check for any
> overlapping flow entries in the requested table. Two flow entries overlap
> if a single packet may match
> both, and both entries have the same priority. If an overlap conflict
> exists between an existing flow entry and the add request, the switch
> must refuse the addition and respond with an ofp_error_msg with
> OFPET_FLOW_MOD_FAILED type and OFPFMFC_OVERLAP code."
> the consequence is that I believe your explanation allows for no error
> if we add exactly the same rule but the spec requires an error. How do
> real switches implement this? Does this affect the validity of your
> algorithms?
> in Algorithm 6 you seems to be assuming that same match with same
> actions won't cause an error but same match different actions will. This
> is neither consistent with your explanation of overlap check nor the spec
> Algorithm 3 I believe commands cmd1 and cmd2 must be either
> bundled together with guaranteed ordering or separated by a barrier,
> otherwise the switch might reorder them and render the cleanup

> Why is Algorithm 8 "more efficient alternative to Algorithm 7" ? It
> looks to be doing much more work

Reviewer 3
==========

> This is a intriguing little article that shows that a recent version of
> OpenFlow (which version?!) has the raw ingredients needed to implement a very general
> transactional interface for updating flow tables. It is obvious that this is
> an accident. The OpenFlow primitives that the article uses actually have a
> very narrow interface, but the paper uses them in a clever way to build very
> general transactions: a transaction is an atomic sequence of flowtable
> writes and assertions. i.e., the writes succeed or fail atomically only if all the
> assertions are true.
> This article is primarily a theoretical exercise and doesn't address
> performance concerns:
> Do real switches have the space to use all these extra rules for
> coordination?
> Do switches actually implement the features of OpenFlow used by this article
> (i.e., check_overlap, bundles, arbitrary bitmasking of the METADATA field,
> etc.) and do they implement them fast?
> Being optimistic, I think that the real value of this article may be to
> show that OpenFlow might as well provide a more general transactional interface.
> Without any empirical evidence, I doubt that the scheme proposed in the article
> will actually work in practice.

> A deeper issue that this article doesn't address is whether this kind of
> abstraction is even desirable. A simpler abstraction may be to have one
> controller for a small set of switches and have the controllers coordinate
> with each other.
> Transactions are a powerful primitive, but apart from the simple load
> balancer in Algorithm 1, the paper doesn't delve into how applications should be
> built to take advantage of the transactional primitive. I think this is what
> the article sorely lacks. Without it, the article just shows off cool hack.

> MINOR
> pg 1 col 1: "perspective of [a] logically centralized"

> pg 1 col 1: "a distributedsystems one: [multiple]"

> pg 1 col 2: perhaps paper article

> pg 1 col 2: "We [propose] to implement our"

> pg 3 col 1: "as we will argue, asupport for more generic" (delete a)

> pg 3 col 2: "a bundle allows to conditionally"

> In Example 1, why not have the two controllers communicate

> Example 2 says bundling and checkoverlap don't help modify. But, Section 2
> says that the semantics of addflow is to modify if a flow with exactly
> the same priority and match exist

EDITOR
======

> Dear authors,
> Thank you for your submission. The reviews for this paper raise several
> important questions and provide good feedback. We hope they will help
> in improving the paper. Beyond addressing the technical questions and
> presentationrelated
> issues mentioned in the reviews, a revised version
> of the paper should include:
> Examples
> showing how transactions can be used by control
> applications. This will make the paper storyline richer.
> Experiments
> with actual switches to give readers some confidence that
> the proposed transactions can be efficiently implemented.
> Regards. Area Editor.