### 2015-08-31 reviews/response-rev1.txt
some updates

Dear Dina, dear reviewers:

thank you very much for your time and good feedback!

We revised our paper according to your suggestions.

In the following, we discuss the most important changes.

We hope that the new version meets your expectations, and are happy to take any further feedback.

With kind regards,

Liron, Petr, Stefan

Reviewer 1
==========

> Overall, the work seems useful and correct.
> My main concern is scope.
> The paper is essentially a note on OpenFlow hacking, and thus might be
> of limited interest to many CCR readers. However, OpenFlow and similar
> SDN approaches have seen sufficient adoption and research interest in
> recent years that there is utility in this paper appearing in a venue with
> broad reach.

Thank you very much for your feedback. Indeed, we hope and believe that
our work is of relevance beyond specific Openflow versions. In our revision,
we aim to put the fundamental aspects and the benefit of inband synchronization
into a broader context. Having that said, in our revision we also report
in more detail on our proof-of-concept implementation based on POX.

> A lesser concern is novelty. With the addition of atomic bundles to
> OpenFlow, it's not completely surprising that the atomic CAS operation
> can be implemented, but there is definite value in having it spelled out
> cleanly.

We agree, and believe that the implementations with and without bundles
is a non-trivial step forward.

> I found the brief argument for inband
> control vs outofband control unconvincing. There are many highperformance
> systems that use consensus, and a full architectural
> comparison of the two approaches to
> SDN is beyond the scope of this paper. I certainly believe that inband
> control is a useful and interesting design choice, but not just because of
> FLP. It also seems almost certain that a full system would require outofband
> coordination between controllers: a full transactional system needs
> application specific ways of resolving conflicts, which will require
> coordination.

We agree. With this paper, we aim to show feasibility of inband synchronization, 
and we offer a simple interface of synchronization primitives. While inband mechanisms
may reduce the round-trip times and reduce the number of elements a protocol relies upon,
further research is needed to fully answer the out-of-band vs inband
debate.

> I also found 2nd example of the limitations of bundles confusing. A
> simple concrete example would be greatly helpful, and strength the
> argument.

We reorganized the paper, and hope that the contributions are now
put into perspective better.

> Typos/corrections:
> p.1 col. 2 para. (Contributions): 2nd "conflicts" => "a conflict"

Done, thanks.

> "does not affect" => "affects neither"

Thanks.

> " " para. 3: "promote" => "propose" (?)

Thanks.

> p. 3 col. 2 para (Configuration space): "provided a total addressing scheme" sentence is confusing.

We changed this part, thank you.

Reviewer 2
==========

> The idea of the paper is really nice and I believe it can be used in
> practice. However, the presentation of the paper needs major work it
> feels to be switching between highlevel
> overview versus lowlevel details is
> background section necessary? Cannot you introduce details like
> bundles, overlap_checking and actual OpenFlow flags progressively and
> on examples?
> maybe start with another example with controllers updating
> overlapping flows. Then explain ability to do overlap_check. Then
> introduce loadbalancing example and the need for external
> synchronization. Transactions/bundles are kindof
> orthogonal and could
> be explained after you state that you want executeatomic{ops} and
> that there is a native support for it in OpenFlow if ops are OF commands.
> Figure 1 time axis is too inconspicous
> when you described execute{op1,...} and executeatomic{
> op1, ...} I was totally confused what is the definition good for. Only later on I
> realized that operations there are actually not OpenFlow operations but
> your custom synchronization primitives. Please, clarify that in the text.
> what does "mask=1...1" mean? Is it a) exact match or b) everything
> wildcarded? From analysing algorithm 5 I believe it is a)

This is a very good suggestion, and in our revision, we restructured the paper. 
We hope that the new structure puts the contribution into perspective better.

> Do the algorithms 24 really need to use value=x.x?
> Cannot it be "0..0 . x" ? Or "1...1 . x"? Or for an arbitrary fixed
> constant "c.x" ? Maybe elaborate in the text on that point because it
> looks like magic to me.
> In any case, you should explain more why "x masked by self1"
> overlaps with "x masked by self2" exactly iff self1 != self2

Thanks, we tried to clarify this.

> On my first reading I missed that claim() operation can actually allow
> more controllers to claim the same value. Please clarify in the text.
> What is the value of such operation? Actually, it looks like a kindof
> "read lock". following the thought, additionally to claim() and unclaim(), could you
> have algorithms for exclusive lock() and unlock()? Maybe even explain it
> sooner than claim() so that people can think about wellknown
> semantic first.

Thank you, we revised this part and tried to clarify.

> There looks to be an inconsistency of what happens with check_overlap
> flag for the exactly same rule:
> You say "a FlowMod command can be equipped with a the check
> overlap flag: if the switch maintains a flow entry with an overlapping but
> not identical match part with the same priority but a different action,
> and if the two rules have the same priority, then FlowMod will fail."
> OF 1.4 spec says "For add requests (OFPFC_ADD) with the
> OFPFF_CHECK_OVERLAP flag set, the switch must first check for any
> overlapping flow entries in the requested table. Two flow entries overlap
> if a single packet may match
> both, and both entries have the same priority. If an overlap conflict
> exists between an existing flow entry and the add request, the switch
> must refuse the addition and respond with an ofp_error_msg with
> OFPET_FLOW_MOD_FAILED type and OFPFMFC_OVERLAP code."
> the consequence is that I believe your explanation allows for no error
> if we add exactly the same rule but the spec requires an error. How do
> real switches implement this? Does this affect the validity of your
> algorithms?

We elaborated more on these technical parts, and also now include a short
discussion of a proof-of-concept implementation.

> in Algorithm 6 you seems to be assuming that same match with same
> actions won't cause an error but same match different actions will. This
> is neither consistent with your explanation of overlap check nor the spec

We agree that this part was confusing, and changed the description now.

> Algorithm 3 I believe commands cmd1 and cmd2 must be either
> bundled together with guaranteed ordering or separated by a barrier,
> otherwise the switch might reorder them and render the cleanup

FIXME

> Why is Algorithm 8 "more efficient alternative to Algorithm 7" ? It
> looks to be doing much more work

Indeed, the algorithm seems to be longer, and we now clarified what we mean by more efficient.

Reviewer 3
==========

> This is a intriguing little article that shows that a recent version of
> OpenFlow (which version?!) has the raw ingredients needed to implement a very general
> transactional interface for updating flow tables. It is obvious that this is
> an accident. The OpenFlow primitives that the article uses actually have a
> very narrow interface, but the paper uses them in a clever way to build very
> general transactions: a transaction is an atomic sequence of flowtable
> writes and assertions. i.e., the writes succeed or fail atomically only if all the
> assertions are true.
> This article is primarily a theoretical exercise and doesn't address
> performance concerns:
> Do real switches have the space to use all these extra rules for
> coordination?
> Do switches actually implement the features of OpenFlow used by this article
> (i.e., check_overlap, bundles, arbitrary bitmasking of the METADATA field,
> etc.) and do they implement them fast?
> Being optimistic, I think that the real value of this article may be to
> show that OpenFlow might as well provide a more general transactional interface.
> Without any empirical evidence, I doubt that the scheme proposed in the article
> will actually work in practice.

We indeed understand our article as a first step, and a proof of concept.
However, much more work is needed to evaluate its performance and scalability,
and we aim to do this in our future research. Nevertheless, our mechanisms
can readily be implemented in today's Openflow protocol. In our revision,
we also report on such an implementation.

> A deeper issue that this article doesn't address is whether this kind of
> abstraction is even desirable. A simpler abstraction may be to have one
> controller for a small set of switches and have the controllers coordinate
> with each other.
> Transactions are a powerful primitive, but apart from the simple load
> balancer in Algorithm 1, the paper doesn't delve into how applications should be
> built to take advantage of the transactional primitive. I think this is what
> the article sorely lacks. Without it, the article just shows off cool hack.

This is a very good point. While it is difficult to prove that our abstraction
is the best or even the right one, we believe that our approach represents a meaningful
first step. Moreover, we believe that our contribution goes beyond an "openflow hack".
By restructuring the paper, we seek to put our contribution into a more general
and broader perspective.

> MINOR
> pg 1 col 1: "perspective of [a] logically centralized"

Thanks, done.

> pg 1 col 1: "a distributedsystems one: [multiple]"

Fixed.

> pg 1 col 2: perhaps paper article

Fixed.

> pg 1 col 2: "We [propose] to implement our"

Fixed.

> pg 3 col 1: "as we will argue, asupport for more generic" (delete a)

Fixed.

> pg 3 col 2: "a bundle allows to conditionally"

Fixed.

> In Example 1, why not have the two controllers communicate
> Example 2 says bundling and checkoverlap don't help modify. But, Section 2
> says that the semantics of addflow is to modify if a flow with exactly
> the same priority and match exist

In our revision, we restructured these parts and hope we found
a omre intuitive way to put this into perspective.

EDITOR
======

> Dear authors,
> Thank you for your submission. The reviews for this paper raise several
> important questions and provide good feedback. We hope they will help
> in improving the paper. Beyond addressing the technical questions and
> presentationrelated
> issues mentioned in the reviews, a revised version
> of the paper should include:
> Examples
> showing how transactions can be used by control
> applications. This will make the paper storyline richer.
> Experiments
> with actual switches to give readers some confidence that
> the proposed transactions can be efficiently implemented.
> Regards. Area Editor.

Thank you for your good comments. In our revision, we restructured
the paper and also elaborate more on the proof of concept implementation,
which shows the feasibility of our approach. Of course, for an extensive
study of the performance and overheads, more work is needed. However,
we hope that our article provides a worthwhile first step and
contribution for the community.